---
title: "Exam2"
author: "Daniel Nguyen"
date: "11/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exam 2

### Setup
```{r}
# Recommended packages
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(modelr)
library(purrr)
library(gridExtra)


# Data 
load('census_city_2009.Rda')
load('census_city_2014.Rda')
load('census_city_2019.Rda')

# Update Column name

colnames(city_2009) <- c("state","place","city","population")
colnames(city_2014) <- c("state","place","city","population")
colnames(city_2019) <- c("state","place","city","population","household_income","housing_units","population_in_house","GEOID","latitude","longitude")

#Combine city data frame 
city <- city_2009 %>% select(city) %>% inner_join(city_2014, by="city") %>% select(city) %>% inner_join(city_2019, by="city")

# Filter
city <- city %>% filter(grepl(" city,",city), population >=25000)


load('city_models.Rda')
cities <- names(models)
city_2029 <- data.frame()
for(i in seq_along(cities)){
  city_temp <- data.frame(city = cities[i], year = 2029)
  city_temp <- city_temp %>% add_predictions(models[[i]],var = 'population')
  city_2029 <- rbind(city_2029, city_temp)
}
colnames(city_2029)[3] <- "population_2029"
city <- city %>% inner_join(city_2029, by='city')

# New calculation column
city <- city %>% mutate(homeless=population - population_in_house,
                        homeless_rate = homeless/population,
                        expected_growth = population_2029-population,
                        expected_growth_rate = expected_growth/population,
                        expected_homeless = homeless_rate * population_2029,
                        expected_homeless_increase = expected_homeless-homeless,
                        household_size = population_in_house/housing_units)


########  Questions ############
#1.What 3 cities have the most homeless in 2019?
## Show 2 tables. One with the highest number of homeless and one with the highest rate of homeless in 2019.

highest_number_homeless <- city %>% arrange(desc(homeless)) %>% slice(1:3) %>% select(city,homeless)
highest_rate_homeless <- city %>% arrange(desc(homeless_rate)) %>% slice(1:3) %>% select(city,homeless_rate)


#2. What are the fastest growing cities? Show 3 plots combined with grid arrange.
top_city <- city %>% mutate(top='Others') %>% arrange(desc(expected_growth))
plot1 <- ggplot(top_city[1:10,], aes(x=city, y=expected_growth)) +geom_bar(stat='identity') + coord_flip() + labs(title="Top 10 cities with expected growth")
top_city[1:10,]$top <- 'Top 10 Growth'
top_city <- top_city %>% arrange(desc(expected_growth_rate))
plot2 <- ggplot(top_city[1:10,], aes(x=city, y=expected_growth_rate)) + labs(title="Top 10 cities with expected growth rate") +geom_bar(stat='identity') + coord_flip()
top_city[1:10,]$top <- 'Top 10 Growth Rate'
top_city$top <- factor(top_city$top, levels = c('Top 10 Growth','Top 10 Growth Rate','Others'))
plot3 <- ggplot(top_city, aes(x=expected_growth, y=expected_growth_rate, color=top)) + geom_point() + scale_colour_manual(values=c('red','green','gray'))+ labs(title="Expected growth and expected growth rate")


#3. Do growing cities have higher homeless rates?
plot4 <- ggplot(top_city, aes(x=expected_growth_rate, y=homeless_rate)) +geom_point()
score <-  cor(top_city$homeless_rate, top_city$expected_growth_rate)

correlation <- top_city %>% keep(is.numeric) %>% select(-year) %>% pivot_longer(names_to='key', values_to ='value', -homeless_rate) %>% group_by(key) %>% summarize(cor=cor(value,homeless_rate,use='complete.obs'))%>% 
  mutate(Rank = dense_rank(1-abs(cor))) %>% arrange(Rank) 

#4.What cities have the highest expected increase in homelessness and how could they eliminate homelessness?

top_city <- top_city %>% mutate(build=ceiling(expected_homeless/household_size))
top_city <- top_city %>% mutate(share=(population_2029/housing_units)-household_size)
top_city <- top_city %>% mutate(top_homeless='Others') %>% arrange(desc(expected_homeless_increase))
top_city[1:100,]$top_homeless <- 'Top Homeless Increase'
top_city$top_homeless <- factor(top_city$top_homeless, levels = c('Top Homeless Increase','Others'))
plot5 <- ggplot(top_city, aes(x=longitude,y=latitude,color=top_homeless)) + geom_point() + borders('state') + scale_colour_manual(values=c('red','green'))
#Average for all cities
house_and_share <- data.frame(avg_build=mean(top_city$build), avg_share=mean(top_city$share))

more_house <- top_city %>% arrange(desc(build)) %>% slice(1:10)
plot5 <- ggplot(more_house, aes(x=city, y=build)) + labs(title="Top 10 cities that needs to build more houses") +geom_bar(stat='identity') + coord_flip()
more_share <- top_city %>% arrange(desc(share)) %>% slice(1:10)
plot6 <- ggplot(more_share, aes(x=city, y=share)) + labs(title="Top 10 cities that needs to increase housing size") +geom_bar(stat='identity') + coord_flip()
```


### Question. 
1. What 3 cities have the most homeless in 2019?


Highest number of homeless: `r kable(highest_number_homeless)`

Highest rate of homeless: `r kable(highest_rate_homeless)`


2. What are the fastest growing cities? Show 3 plots combined with grid arrange.

```{r}
grid.arrange(plot1,plot2,plot3,ncol=1)
```

3. Do growing cities have higher homeless rates?

```{r}
plot4
```

Correlation score: `r score`

Based on the score and the plot, it shows that they have a very slight negative correlation

How about other features?

`r kable(correlation)`

Since others relate to homeless, I would pick household_income and household_size. They have a negative correlation score, meaning that cities which have small household size and less income will likely to have higher rate of homeless.


4. What cities have the highest expected increase in homelessness and how could they eliminate homelessness?

Average share and build for all cities:
`r kable(house_and_share)`

Top 10 cities that need more houses or increasing the housing size to eliminate homeless in 2029

```{r}
grid.arrange(plot5,plot6)
```